
[data from here](https://infogram.com/1p7ve7kjeld1pebz2nm0vpqv7nsnp92jn2x?live)

imports and libraries
```{r}
library(tidyverse)
cases <- read.csv("cases_daily_aus.csv") 
colnames(cases) <- c("date", "new.cases")
#add 2020 to all date strings
cases <- mutate(cases, date = paste(date,"-2020", sep = ""))
#format date strings as date objects
cases <- mutate(cases, date = as.Date(date, format = "%e-%b-%Y"))
#use weekdays() to identify weekday
cases <- mutate(cases, weekday = weekdays(date))
cases
```
Frequency per weekday
```{r}
frequency.by.weekday <- cases %>% group_by(weekday) %>% summarise( sum = sum(new.cases))
frequency.by.weekday <- merge(frequency.by.weekday, count(cases, weekday))
#rename columns
colnames(frequency.by.weekday) <- c("weekday","sum.new.cases","count.of.weekday")
#set weekday as factor with Sunday as start of week
frequency.by.weekday$weekday <- factor(frequency.by.weekday$weekday, levels = 
      c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))
#reorder so that data.frame follows this order
frequency.by.weekday <- frequency.by.weekday[order(frequency.by.weekday$weekday),]
#print!
frequency.by.weekday
```
```{r}
ggplot(frequency.by.weekday, aes(weekday, sum.new.cases)) +
  geom_col()
```
Distribution of cases per day
```{r}
hist(cases$new.cases)
```
Cases below and above 50
```{r}
cases.binned.weekday <- cases %>% filter(new.cases < 50) %>%
  group_by(weekday) %>% count(weekday)
colnames(cases.binned.weekday) <- c("weekday","under.50")

cases.binned.weekday <- cases %>% filter(new.cases > 50) %>%
  group_by(weekday) %>% count(weekday) %>%
  merge(cases.binned.weekday, .)

colnames(cases.binned.weekday) <- c("weekday","under.50", "over.50")

cases.binned.weekday
```
Is bias due to a few very big case days? They absolutely bias the results. Unsure how to account for this yet.
```{r}
saturday.cases <- cases %>% filter(weekday == "Saturday")
hist(saturday.cases$new.cases, breaks = dim(saturday.cases)[1])
```
Probably a better way to do this, because this way I have to adjust by an off-by-one error. First generate slopes using lm() for each 7 day range, sunday to saturday. Put into vector `slopes`. Then calculate start dates from cases$date using seq(), make sure first day is 26-Jan since this is the first day included in the slopes calculation.
```{r}
slopes <- c()
for (i in seq(2,dim(cases)[1],7)) {
  index_7 <- i + 6
  temp.df <- cases[i:index_7,]
  temp.df <- drop_na(temp.df) #last entry might have NA rows
  slopes <- append(slopes, coefficients(lm(new.cases~date, temp.df))['date'])
}
dates <- cases$date[1] + seq(1,dim(cases)[1]-1,7) #off by one error here, minus by 1 to get right date
```
Maybe should try graphing this.

Will need to normalise so that each graph has same magnitude. I think if I take min(x) for vector x of new.case counts for each 7 day interval it should at least be alright.

For the weekday with minimum new.cases per 7 day interval, there are sometimes multiple weekdays. In the seven.day stats table they are assigned "multiple". I will make a separate table to illustrate all of them. Note that most multiple day counts come from early in the outbreak when there were many zero days.

```{r}
mins = c()
means = c()
maxs = c()
min_weekday <- c()
max_weekday <- c()
n_days <- c()
slopes <- c()

for (i in seq(2,dim(cases)[1],7)) {
  #select an interval of seven days from Sunday to Saturday
  index_7 <- i + 6
  temp.df <- cases[i:index_7,]
  temp.df <- drop_na(temp.df) #last entry might have NA rows
  
  #calculate a linear regression slope for each interval
  slopes <- append(slopes, coefficients(lm(new.cases~date, temp.df))['date'])
  
  # get mins and means
  mins <- append(mins, min(temp.df$new.cases))
  means <- append(means, mean(temp.df$new.cases))
  maxs <- append(maxs, max(temp.df$new.cases))
  
  #append unique weekdays which have min value
  min_weekday_temp <- filter(temp.df, new.cases == min(temp.df$new.cases))$weekday
  if (length(min_weekday_temp) > 1) {
    min_weekday_temp <- c("multiple")
  }
  min_weekday <- append(min_weekday, min_weekday_temp)
  
  #append unique weekdays which have max value
  max_weekday_temp <- filter(temp.df, new.cases == max(temp.df$new.cases))$weekday
  if (length(max_weekday_temp) > 1) {
    max_weekday_temp <- c("multiple")
  }
  max_weekday <- append(max_weekday, max_weekday_temp)
  
  #get number of days in each interval for debugging purposes
  n_days <- append(n_days, dim(temp.df)[1])
}
seven.day <- data.frame(date = dates, n_days = n_days, slopes = slopes, 
                        mins = mins, means=means, maxs = maxs, 
                        min_weekday = min_weekday, max_weekday = max_weekday)
seven.day
```
I am unsure if min value or mean value makes a better normalising value here. Certainly typical loss functions use mean value so maybe that makes more sense? I am not sure.

A count of which day has the minimum new cases count for a given 7 day interval shows most days are Sundays. But that this is not always the case. This includes the many weeks where there are multiple days with the minimum number of cases (usually zero) for early in the outbreak.
```{r}
min_weekday <- c()
for (i in seq(2,dim(cases)[1],7)) {
  index_7 <- i + 6
  temp.df <- cases[i:index_7,]
  temp.df <- drop_na(temp.df) #last entry might have NA rows
  
  min_weekday_temp <- filter(temp.df, new.cases == min(temp.df$new.cases))$weekday
  min_weekday <- append(min_weekday, min_weekday_temp)
}
table(min_weekday)
```
Masking out weeks with multiples doesn't change things too much, but all the fridays disappear:
```{r}
table(seven.day$min_weekday)
```
We can naively assume that for most weeks, Sunday will have the lowest reported number of cases, and Friday is least likely to report the min number of cases.
Since there are few weeks with multiple max days, we can use a much simpler code block to get those numbers:
```{r}
table(seven.day$max_weekday)
```
Unmasking multiples gets a similar story:
```{r}
max_weekday <- c()
for (i in seq(2,dim(cases)[1],7)) {
  index_7 <- i + 6
  temp.df <- cases[i:index_7,]
  temp.df <- drop_na(temp.df) #last entry might have NA rows
  
  max_weekday_temp <- filter(temp.df, new.cases == max(temp.df$new.cases))$weekday
  max_weekday <- append(max_weekday, max_weekday_temp)
}
table(max_weekday)
```

```{r}
cases
```
```{r}
cases$ordinal <- 1:dim(cases)[1]
cases %>% group_by(ordinal = cut(ordinal, breaks = dim(cases)[1]/7)) %>%
  summarise_each(funs(mean))
```
```{r}
cases %>% group_by(ordinal = cut(ordinal, breaks = dim(cases)[1]/7)) %>%
  summarise(mean(new.cases), min(new.cases), max(new.cases), n_distinct(weekday))
```
```{r}
cases
```






























